---
title: "Persuasive Privacy"
subtitle: "ICML 2026"
author: "Joshua J Bon$^\\dagger$ James Bailie$^\\ast$ Judith Rousseau$^\\diamond$ Christian P Robert${^\\diamond}{^\\circ}$"
institute: "$^\\dagger$Adelaide University $^\\ast$Chalmers University $^\\diamond$Université Paris–Dauphine PSL $^\\circ$University of Warwick"
bibliography: refs.bib
format:
  revealjs:
    html-math-method: katex
    theme: default
    df-print: paged
    incremental: true 
    css: style.css
    chalkboard: true
editor: source
---

## Persuasive Privacy

::: {.fragment}

A **semantics-first** understanding of data privacy where

:::

- Privacy definitions are systematically constructed from an **agent-based game**
- Game assumptions have real-world interpretations
- Probabilistic and pure **DP special cases**
- Rényi and $f$-divergence DP also by changing assumptions
- Privacy developed for **deterministic mechanisms**

## A Game of Persuasion

::: {style="color:blue;"}
Sender chooses a mechanism that **persuades** Receiver to make decisions limiting privacy loss.
:::

:::: {.columns}

::: {.column width="50%"}
### Sender

_Designs mechanism_

_Worst-case analysis in_

:::{.nonincremental}
- _data value_

- _Receiver's data-prior_
:::
:::

::: {.column width="50%"}
### Receiver

_Observes mechanism output_ 

_Minimizes expected loss w.r.t. data-posterior,_

_Bayesian agent^[Similar to Bayesian persuasion [@kamenica2011bayesian]]_
:::

::::



## Data privacy game

```{r dtp, engine = 'tikz',   fig.align = 'center', cache=TRUE}

\begin{tikzpicture}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{fit}
\usetikzlibrary{shapes}

\tikzstyle{st} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{ad} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{pb} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{pbmech} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{bayes} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{desc} = [rectangle, minimum width=2.5cm, rounded corners, minimum height=0.6cm,text centered, fill=gray!10]
\tikzstyle{greybox} = [draw=gray, rounded corners, inner sep=0.5cm, thick]
\tikzstyle{publicbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=teal!30]
\tikzstyle{privatebox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=purple!30]

% Statistician
\node (sdata) [st] {$x$};
\node (sdatad) [desc, left=0pt of sdata] {Data};
\node (mech) [pbmech, below=0.5cm of sdata] {$M$};
\node (mechd) [desc, left=0pt of mech] {Mechanism};
\node (sstat) [pb, below=0.5cm of mech] {$T$};
\node (sstatd) [desc, left=0pt of sstat] {Statistic};
\node (spriv) [st, below=1cm of sstat] {$\rho(d,x)$};
\node (sprivd) [desc, left=0pt of spriv] {Privacy};

% Adversary
\node (adata) [ad, right=3cm of sdata] {$Q$};
\node (adatad) [desc, right=0pt of adata] {Data-prior};
\node (apost) [ad, right=3cm of sstat] {$Q_T$};
\node (apostd) [desc, right=0pt of apost] {Data-posterior};
\node (aact) [ad, below=1cm of apost] {$d$};
\node (aactd) [desc, right=0pt of aact] {Decision};
\node (aloss) [ad, below=1cm of aact] {$\ell$};
\node (alossd) [desc, right=0pt of aloss] {Loss};

% info
\node (public) [publicbox, below=1.5cm of sprivd, xshift=-1.5cm] {};
\node (publicd) [desc, right=0pt of public] {Shared};
\node (private) [privatebox, right=0.5cm of publicd] {};
\node (privated) [desc, right=0pt of private] {Private};


% Statistician
\draw[thick, -] (sdata) -- (mech);
\draw[thick, -Latex] (mech) -- (sstat);


% Adversary
\draw[thick, -Latex] (adata) -- (apost);
\draw[thick, -Latex] (apost) -- (aact);
\draw[thick, -Latex] (aloss) -- (aact);

%Cross
\draw[thick, -Latex] (mech) -- (apost);
\draw[thick, -Latex] (sstat) -- (apost);
\draw[thick, -Latex] (aact) -- (spriv);

% boxes
\node[greybox, fit=(sdata) (sdatad) (sstat) (sstatd) (spriv) (sprivd)] (statbox) {};
\node[greybox, fit=(adata) (adatad) (apost) (apostd) (aact) (aactd) (aloss) (alossd)] (advbox) {};

\node[anchor=south east, xshift=-0.9cm] at (statbox.north) {Sender};
\node[anchor=south east] at (advbox.north) {Receiver model};

\end{tikzpicture}

```

## Game under persuasive privacy

```{r dtp3, engine = 'tikz',   fig.align = 'center', cache=TRUE}

\begin{tikzpicture}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{fit}
\usetikzlibrary{shapes}

\tikzstyle{st} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{ad} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{pb} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{as} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=orange!30]
\tikzstyle{im} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=blue!30]
\tikzstyle{pbmech} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{bayes} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{desc} = [rectangle, minimum width=2.5cm, rounded corners, minimum height=0.6cm,text centered, fill=gray!10]
\tikzstyle{greybox} = [draw=gray, rounded corners, inner sep=0.5cm, thick]
\tikzstyle{publicbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=teal!30]
\tikzstyle{privatebox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=purple!30]
\tikzstyle{assbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=orange!30]
\tikzstyle{impbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=blue!30]

% Statistician
\node (sdata) [st] {$x$};
\node (sdatad) [desc, left=0pt of sdata] {Data};
\node (mech) [pbmech, below=0.5cm of sdata] {$M$};
\node (mechd) [desc, left=0pt of mech] {Mechanism};
\node (sstat) [pb, below=0.5cm of mech] {$T$};
\node (sstatd) [desc, left=0pt of sstat] {Statistic};
\node (spriv) [st, below=1cm of sstat] {$\rho(d,x)$};
\node (sprivd) [desc, left=0pt of spriv] {Privacy};

% Adversary
\node (adata) [as, right=3cm of sdata] {$Q\in\mathcal{Q}_x$};
\node (adatad) [desc, right=0pt of adata] {Data-prior};
\node (apost) [im, right=3cm of sstat] {$Q_T$};
\node (apostd) [desc, right=0pt of apost] {Data-posterior};
\node (aact) [im, below=1cm of apost] {$d$};
\node (aactd) [desc, right=0pt of aact] {Decision};
\node (aloss) [as, below=1cm of aact] {$\ell = \rho$};
\node (alossd) [desc, right=0pt of aloss] {Loss};

% info
\node (public) [publicbox, below=1.25cm of sprivd, xshift=-1.5cm] {};
\node (publicd) [desc, right=0pt of public] {Shared};
\node (private) [privatebox, right=0.5cm of publicd] {};
\node (privated) [desc, right=0pt of private] {Private};
\node (ass) [assbox, below=0.25cm of public] {};
\node (assd) [desc, right=0pt of ass] {Assumed};
\node (imp) [impbox, right=0.5cm of assd] {};
\node (impd) [desc, right=0pt of imp] {Implied};


% Statistician
\draw[thick, -] (sdata) -- (mech);
\draw[thick, -Latex] (mech) -- (sstat);


% Adversary
\draw[thick, -Latex] (adata) -- (apost);
\draw[thick, -Latex] (apost) -- (aact);
\draw[thick, -Latex] (aloss) -- (aact);

%Cross
\draw[thick, -Latex] (mech) -- (apost);
\draw[thick, -Latex] (sstat) -- (apost);
\draw[thick, -Latex] (aact) -- (spriv);

% boxes
\node[greybox, fit=(sdata) (sdatad) (sstat) (sstatd) (spriv) (sprivd)] (statbox) {};
\node[greybox, fit=(adata) (adatad) (apost) (apostd) (aact) (aactd) (aloss) (alossd)] (advbox) {};

\node[anchor=south east, xshift=-0.9cm] at (statbox.north) {Sender};
\node[anchor=south east, xshift=2.35cm] at (advbox.north) {Receiver model + assumptions};

\end{tikzpicture}

```

## Persuasive Privacy Paper

- New interpretations of post-processing property

- Properties: **composition** and **Receiver post-processing**

- Deterministic mechanisms: empirical mean, cell suppression

- Recovery of existing DP privacy definitions

- arXiv: [https://arxiv.org/abs/2601.22945](https://arxiv.org/abs/2601.22945)


## References

::: {#refs}
:::



