---
title: "Persuasive Privacy"
subtitle: "University of Melbourne, 23^rd^ March 2026"
author: "Joshua J Bon"
institute: "Adelaide University"
bibliography: refs.bib
format:
  revealjs:
    html-math-method: katex
    theme: default
    df-print: paged
    incremental: true 
    css: style.css
    chalkboard: true
editor: source
---

## My research

```{dot}
graph G {
  layout=neato;
  subgraph main {
    node [fixedsize=true, width=0.7, height=0.4];
    SBI -- SMC;
    SMC -- Privacy;
    Privacy -- SBI;
  }
  
}
```

## My research

:::: {.columns}

::: {.column width="60%"}

::: {.nonincremental}

**Statistics**

- Bayesian modelling[^1] [^2]
- Efficient computation[^3] [^4] [^5]
- Approximate models^3^ ^4^

:::

:::

::: {.column width="40%"}

```{dot}
graph G {
  layout=neato;
  subgraph main {
    node [fixedsize=true, width=0.7, height=0.4];
    SBI -- SMC;
  }
  
}
```

:::

::::

[^1]: @oflaherty2021teen _Demography_
[^2]: @bon2019polling _JRSS A_
[^3]: @bon2021dasmc _Stat & Comp_
[^4]: @bon2025cal _JMLR_
[^5]: @bon2025rbsmc _arXiv_

## My research

:::: {.columns}

::: {.column width="60%"}

::: {.nonincremental}

**Privacy**

- Federated & transfer learning[^6] [^7]
- Private (Bayesian) computation[^8]
- Approximate models[^9]

:::

:::

::: {.column width="40%"}

```{dot}
graph G {
  layout=neato;
  subgraph main {
    node [fixedsize=true, width=0.7, height=0.4];
    Privacy -- SMC;
    Privacy -- SBI;
  }
  
}
```

:::

::::

[^6]: @hassan2024federated _arXiv_
[^7]: @bretherton2026principled _Bayesian Analysis_
[^8]: @bon2026priv _arXiv_
[^9]: @bon2020acld _J Official Statistics_

## My research

```{dot}
graph G {
  layout=neato;
  subgraph main {
    node [fixedsize=true, width=0.7, height=0.4];
    SBI;
    SMC;
    Privacy;
    SBI -- SMC;
    SMC -- Privacy;
    Privacy -- SBI;
  }
  
}
```

## Today's talk

```{dot}
graph G {
  layout=neato;
  subgraph main {
    node [fixedsize=true, width=0.7, height=0.4];
    SBI;
    SMC;
    Privacy [color="red"];
    SBI -- SMC;
    SMC -- Privacy;
    Privacy -- SBI;
  }
  
}
```

## Today's talk

:::{.nonincremental}

1. Differential privacy
2. Privacy from the ground up
3. Persuasive privacy
4. Differential privacy redux
5. Privacy without noise

:::


See: **Bon**, Bailie, Rousseau & Robert (2026) arXiv 2601.22945 

_Under review at ICML 2026_


# Differential privacy

## DP framework

$$\text{Dataset}~x \overset{\text{release}}{\longrightarrow} \text{statistic}~T$$

- $T \sim M(x,\cdot)$, with some mechanism $M$
- $T = \frac{1}{n}\sum_{i=1}^{n} x_i + Z$ with $Z \sim \mathcal{N}(0,\sigma^2)$

::: {.fragment}
$$M(x,\cdot) = \mathcal{N}(\overline{x},\sigma^2)$$
:::

## DP Intuition

- To limit information leaked about data $x$ by statistic $T$,
- Control the sensitivity of the mechanism $M(x,\cdot)$ w.r.t. $x$.

::: {.fragment}

PLOTS HERE

:::

## DP Definition

Borrow from Lipschitz continuity [@bailie2024general].

::: {.fragment}
$$d_\text{P}[M(x,\cdot),M(x^\prime,\cdot)] \leq \epsilon~d_\text{X}[x,x^\prime]$$
:::

::: {.fragment}
for all $x,x^\prime \in \mathcal{D}$.
:::

## DP Definition

A mechanism $M$ is $\epsilon$-DP if

::: {.fragment}
$$\sup_{S\in\mathcal{S}}\left\vert\ln\frac{M(x,S)}{M(x^\prime,S)}\right\vert \leq \epsilon$$
:::

::: {.fragment}
for all $x,x^\prime \in \mathcal{D}$, differing by at most one element.
:::

::: {.fragment}

:::{.callout-note}
## Pure Differential Privacy 

$$M(x,S) \leq \exp\{\epsilon\} M(x^\prime, S)$$
for all $x \sim x^\prime$ and $S \in \mathcal{S}$.

:::
:::

## Some DP Properties

Focusing on (pure) $\epsilon$-DP.

1. **Composition**: If $M_1$ and $M_2$ are $\epsilon$-DP then

::: {.fragment}

$$M_1 \otimes M_2 ~\text{is}~ 2\epsilon\text{-DP}$$

:::

2. **Post-processing**: If $K$ is independent of the data then
 
::: {.fragment}

$$M_1 K ~\text{is}~ \epsilon\text{-DP}$$

:::

## Some DP Properties

Focusing on $\epsilon$-DP. Examples:

:::{.nonincremental}

1. **Composition**:

If releasing the mean and variance of a dataset is $\epsilon$-DP respectively, then the joint release is $2\epsilon$-DP.

2. **Post-processing**:

If $M$ constructs a histogram that is $\epsilon$-DP, then any quantile from the histogram will also be $\epsilon$-DP.

:::

## Variants of DP

There are many variants and relaxations of DP.

- Probabilistic DP [@guingona2023comparing]

::: {.fragment}
$$\inf_{x\sim x^\prime}\mathbb{P}_{x}\left[ m(x,T) \leq \exp\{\epsilon\} m(x^\prime, T) \right] \geq 1 - \delta$$
:::

- Approximate DP [@dwork2006approxdp]

::: {.fragment}
$$m(x,S) \leq \exp\{\epsilon\} m(x^\prime, S)  + \delta$$
for all $S \in \mathcal{S}$.
:::

## Variants of DP

There are many variants and relaxations of DP.

- Rényi DP
- Gaussian DP
- Concentrated DP
- Many more!

:::{.fragment}
All related to Lipschitz interpretation 
:::

## Pure DP Semantics

- Hypothesis testing
  - Type I & II error controlled by $\epsilon$ [@wasserman2010statistical]
- Bayesian evidence
  - Prior-to-posterior ratio controlled [@dwork2006calibrating]
  - Posterior-to-posterior ratio controlled [@kasiviswanathan2014semantics]
- Game theory 
  - Controls expected utility [@zhang2021more]

## Pure DP Semantics

::: {.nonincremental}
- Hypothesis testing
- Bayesian evidence
- Game theory
:::

::: {.fragment}
DP Semantics are... 
:::

::: {.fragment}
**... post-hoc explanations of DP**
:::

::: {.fragment}
and motivated to find privacy definitions that are relevant for deterministic mechanisms. 
:::

# Privacy from the ground up

## Overview of statistical data privacy

:::{.callout-tip}
## Sender

:::{.fragment}
Government agency release **small area statistics** on a disease
:::

:::

:::{.callout-warning}
## Receiver

:::{.fragment}
Insurance company **raises premiums** using inferred disease prevalence in a small town
:::

:::

:::{.callout-important}
## Effect on data privacy 

:::{.fragment}
Statistic release $\longrightarrow$ adversarial decisions $\longrightarrow$ privacy effect
:::

:::

# Anatomy of statistical data privacy

## The players

### Sender

- Custodian of data
- Will use mechanism to release information about data

### Receiver

- Makes decision based on revealed information
- Decision affects Sender's privacy

## Privacy function

Sender's privacy measured with a **privacy function**

$$\rho:\mathcal{D} \times \mathsf{X} \rightarrow \mathbb{R}$$

- Receiver's decision $\quad d_i \in \mathcal{D}$

- Data value $\quad x \in \mathsf{X}$

:::{.fragment}

$\rho$ orders preferences of decisions for a given dataset
:::


- If $d_1$ is preferred to $d_2$ then $\rho(d_1,x) > \rho(d_2,x)$


## Example 1: Privacy function

For some fixed $\kappa>0$,

$$\rho(d,x) = \begin{cases}
  0, & \text{if } x \in d, \vert d \vert < \kappa\\
  1, & \text{otherwise},
\end{cases}$$


:::{.nonincremental}

- $x \in \mathsf{X} = \mathbb{R}$

- $d \in \mathcal{D} = \{[a,b]:a,b\in \mathbb{R}, a\leq b\}$
:::


## Example 2: Privacy function

<br>

$$\rho(d,x) = -\log d(x)$$

:::{.nonincremental}

- $x \in \mathsf{X} = \mathbb{R}$

- $d \in \mathcal{D} = \{\text{probability density functions on } \mathbb{R}\}$
:::


## The statistic
 
A statistic $T \in \mathsf{T}$ is output from a mechanism $M$ given $x$

$$T \sim M(~\cdot \mid x)$$

- $M:\mathcal{T} \times \mathsf{X} \rightarrow [0,1]$ for $(\mathsf{T},\mathcal{T})$ measurable space

## The statistic

$M$ can be deterministic or randomised

- Summary statistics $\quad\tau:\mathsf{X}\rightarrow\mathsf{T}$
  - $M(~\cdot \mid x) = \delta_{\tau(x)}(\cdot)$ 

- Noisy summary statistics
  - $M(~\cdot \mid x) = \mathcal{N}(~\cdot \mid  \tau(x), \sigma^2)$
  
- Composition of multiple mechanisms

## The statistic

$M$ can be deterministic or randomised

- Result of optimisation
  - MLE: $\quad M(~\cdot \mid x) = \delta_{\tau(x)}(\cdot)$
  - $\tau(x) = \arg\max_{\theta \in \Theta} \log L(x \mid \theta)$
  
- Monte Carlo output
  - MCMC: $\quad M(\mathrm{d}t_{1:N} \mid x) = \prod_{i=1}^N K(\mathrm{d}t_{i} \mid t_{i-1}, x)$

- Noisy optimisation
  - SGD: $\quad M(\mathrm{d}t_N \mid x) = \int_{\mathsf{T}_{-n}}\prod_{i=1}^N K(\mathrm{d}t_{i} \mid t_{i-1}, x)$

## Transperancy

**Privacy class**: A set of mechanisms that satisfy a given privacy definition.

:::{.fragment}
If $\mathfrak{C}$ is the privacy class generated by "$\text{PD}$" then 

$$M \in \mathfrak{C} \Longleftrightarrow M~\text{satisfies}~\text{PD}$$

:::

:::{.fragment}

:::{.callout-note}
## Assumption 1: Transperancy

Sender shares the mechanism $M$ and privacy class $\mathfrak{C}$, for which $M \in \mathfrak{C}$, with Receiver. Further, the
definitions of $M$ and $\mathfrak{C}$ do not depend on the data.

:::

:::

## Receiver

:::{.callout-note}
## Assumption 2: Bayesian adversary
Receiver makes Bayesian decisions
:::

1. Holds a prior distribution over the data $Q$
2. Constructs a posterior distribution $Q_T$ from $(M,T)$
3. Has loss function $\ell: \mathcal{D} \times \mathsf{X} \rightarrow \mathbb{R}$

:::{.fragment}
$\mathfrak{C}$ _may effect Receiver's data posterior. Here assumed not to._
:::

## Anatomy of statistical data privacy

```{r dtp, engine = 'tikz',   fig.align = 'center', cache=TRUE}

\begin{tikzpicture}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{fit}
\usetikzlibrary{shapes}

\tikzstyle{st} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{ad} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{pb} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{pbmech} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{bayes} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{desc} = [rectangle, minimum width=2.5cm, rounded corners, minimum height=0.6cm,text centered, fill=gray!10]
\tikzstyle{greybox} = [draw=gray, rounded corners, inner sep=0.5cm, thick]
\tikzstyle{publicbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=teal!30]
\tikzstyle{privatebox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=purple!30]

% Statistician
\node (sdata) [st] {$x$};
\node (sdatad) [desc, left=0pt of sdata] {Data};
\node (mech) [pbmech, below=0.5cm of sdata] {$M$};
\node (mechd) [desc, left=0pt of mech] {Mechanism};
\node (sstat) [pb, below=0.5cm of mech] {$T$};
\node (sstatd) [desc, left=0pt of sstat] {Statistic};
\node (spriv) [st, below=1cm of sstat] {$\rho(d,x)$};
\node (sprivd) [desc, left=0pt of spriv] {Privacy};

% Adversary
\node (adata) [ad, right=3cm of sdata] {$Q$};
\node (adatad) [desc, right=0pt of adata] {Data prior};
\node (apost) [ad, right=3cm of sstat] {$Q_T$};
\node (apostd) [desc, right=0pt of apost] {Data posterior};
\node (aact) [ad, below=1cm of apost] {$d$};
\node (aactd) [desc, right=0pt of aact] {Decision};
\node (aloss) [ad, below=1cm of aact] {$\ell$};
\node (alossd) [desc, right=0pt of aloss] {Loss};

% info
\node (public) [publicbox, below=1.5cm of sprivd, xshift=-1.5cm] {};
\node (publicd) [desc, right=0pt of public] {Shared};
\node (private) [privatebox, right=0.5cm of publicd] {};
\node (privated) [desc, right=0pt of private] {Private};


% Statistician
\draw[thick, -] (sdata) -- (mech);
\draw[thick, -Latex] (mech) -- (sstat);


% Adversary
\draw[thick, -Latex] (adata) -- (apost);
\draw[thick, -Latex] (apost) -- (aact);
\draw[thick, -Latex] (aloss) -- (aact);

%Cross
\draw[thick, -Latex] (mech) -- (apost);
\draw[thick, -Latex] (sstat) -- (apost);
\draw[thick, -Latex] (aact) -- (spriv);

% boxes
\node[greybox, fit=(sdata) (sdatad) (sstat) (sstatd) (spriv) (sprivd)] (statbox) {};
\node[greybox, fit=(adata) (adatad) (apost) (apostd) (aact) (aactd) (aloss) (alossd)] (advbox) {};

\node[anchor=south east, xshift=-0.9cm] at (statbox.north) {Sender};
\node[anchor=south east] at (advbox.north) {Receiver model};

\end{tikzpicture}

```

## Receiver's optimal decision

From Assumption 1:

:::{.callout-note}
## Receiver's optimal decision
$$
d^{P} \in \arg\inf_{d \in \mathcal{D}} \mathbb{E}_{z \sim Q_T}[ \ell(d,z)]
$$

:::

::: {.fragment}
To model Receiver's decision:
:::

::: {.fragment}
**We need assumptions on $Q$ and $\ell$**
:::

## Receiver model

:::{.callout-note}
## Assumption 3: Adversarial loss function
Receiver has loss function $\ell(d,x) = \rho(d,x)$
:::

- Interpretation 1: Receiver targets privacy

:::{.fragment}
If $d^P_{\ell^\prime}$ is Receiver's optimal decision under $(P,\ell^\prime)$ then
:::

::: {.fragment}
$$\mathbb{E}_{x \sim P}[ \rho(d^{P}_\rho,x)] \leq \mathbb{E}_{x \sim P}[ \rho(d^P_{\ell^\prime},x)]$$
:::


- Interpretation 2: data-averaged **worst-case** for Sender

## Receiver model

:::{.callout-note}
## Assumption 4: Adversarial prior class
Let the data-prior $Q \in \mathcal{Q}_x$
:::

:::{.fragment}
Implies privacy outcome in terms of
:::

:::{.fragment}
:::{.callout-note}
## Receiver's decision
$$
d^{Q_T} \in \arg\inf_{d \in \mathcal{D}} \mathbb{E}_{z \sim Q_T}[ \rho(d,z)]
$$
:::
:::

:::{.fragment}
:::{.callout-note}
## Privacy outcome
$$
\inf_{Q \in \mathcal{Q}_x} \rho(d^{Q_T},x)
$$
:::
:::


## Anatomy of statistical data privacy

```{r dtp5, engine = 'tikz',   fig.align = 'center', cache=TRUE, engine.opts = list(template = "tikz2pdf.tex")}

\begin{tikzpicture}

\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{fit}
\usetikzlibrary{shapes}

\tikzstyle{st} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{ad} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{pb} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{as} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=orange!30]
\tikzstyle{im} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=blue!30]
\tikzstyle{pbmech} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{bayes} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{desc} = [rectangle, minimum width=2.5cm, rounded corners, minimum height=0.6cm,text centered, fill=gray!10]
\tikzstyle{greybox} = [draw=gray, rounded corners, inner sep=0.5cm, thick]
\tikzstyle{publicbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=teal!30]
\tikzstyle{privatebox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=purple!30]
\tikzstyle{assbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=orange!30]
\tikzstyle{impbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=blue!30]

% Statistician
\node (sdata) [st] {$x$};
\node (sdatad) [desc, left=0pt of sdata] {Data};
\node (mech) [pbmech, below=0.5cm of sdata] {$M$};
\node (mechd) [desc, left=0pt of mech] {Mechanism};
\node (sstat) [pb, below=0.5cm of mech] {$T$};
\node (sstatd) [desc, left=0pt of sstat] {Statistic};

\node (strans) [pb, below=0.5cm of sstat] {$\mathfrak{C}$};
\node (stransd) [desc, left=0pt of strans] {Privacy class};

\node (spriv) [st, below=1.5cm of strans] {$\rho(d,x)$};
\node (sprivd) [desc, left=0pt of spriv] {Privacy};

% Adversary
\node (adata) [as, right=3cm of sdata] {$Q\in\mathcal{Q}_x$};
\node (adatad) [desc, right=0pt of adata] {Data prior};
\node (apost) [im, right=3cm of sstat] {$Q_T$};
\node (apostd) [desc, right=0pt of apost] {Data posterior};
\node (aact) [im, below=1cm of apost] {$d$};
\node (aactd) [desc, right=0pt of aact] {Decision};
\node (aloss) [as, below=1cm of aact] {$\ell = \rho$};
\node (alossd) [desc, right=0pt of aloss] {Loss};


% Statistician
\draw[thick, -] (sdata) -- (mech);
\draw[thick, -Latex] (mech) -- (sstat);


% Adversary
\draw[thick, -Latex] (adata) -- (apost);
\draw[thick, -Latex] (apost) -- (aact);
\draw[thick, -Latex] (aloss) -- (aact);

%Cross
\draw[thick, -Latex] (mech) -- (apost);
\draw[thick, -Latex] (sstat) -- (apost);
\draw[thick, -Latex] (aact) -- (spriv);
\draw[thick, -Latex] (strans) -- (apost);

% boxes
\node[greybox, fit=(sdata) (sdatad) (sstat) (sstatd) (spriv) (sprivd) (strans) (stransd)] (statbox) {};
\node[greybox, fit=(adata) (adatad) (apost) (apostd) (aact) (aactd) (aloss) (alossd)] (advbox) {};

\node[anchor=south east, xshift=-0.9cm] at (statbox.north) {Sender};
\node[anchor=south east, xshift=2.35cm] at (advbox.north) {Receiver model + assumptions};

\end{tikzpicture}

```


## Anatomy of statistical data privacy

```{r dtp3, engine = 'tikz',   fig.align = 'center', cache=TRUE}

\begin{tikzpicture}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{fit}
\usetikzlibrary{shapes}

\tikzstyle{st} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{ad} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{pb} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{as} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=orange!30]
\tikzstyle{im} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=blue!30]
\tikzstyle{pbmech} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{bayes} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{desc} = [rectangle, minimum width=2.5cm, rounded corners, minimum height=0.6cm,text centered, fill=gray!10]
\tikzstyle{greybox} = [draw=gray, rounded corners, inner sep=0.5cm, thick]
\tikzstyle{publicbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=teal!30]
\tikzstyle{privatebox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=purple!30]
\tikzstyle{assbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=orange!30]
\tikzstyle{impbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=blue!30]

% Statistician
\node (sdata) [st] {$x$};
\node (sdatad) [desc, left=0pt of sdata] {Data};
\node (mech) [pbmech, below=0.5cm of sdata] {$M$};
\node (mechd) [desc, left=0pt of mech] {Mechanism};
\node (sstat) [pb, below=0.5cm of mech] {$T$};
\node (sstatd) [desc, left=0pt of sstat] {Statistic};
\node (spriv) [st, below=1cm of sstat] {$\rho(d,x)$};
\node (sprivd) [desc, left=0pt of spriv] {Privacy};

% Adversary
\node (adata) [as, right=3cm of sdata] {$Q\in\mathcal{Q}_x$};
\node (adatad) [desc, right=0pt of adata] {Data prior};
\node (apost) [im, right=3cm of sstat] {$Q_T$};
\node (apostd) [desc, right=0pt of apost] {Data posterior};
\node (aact) [im, below=1cm of apost] {$d$};
\node (aactd) [desc, right=0pt of aact] {Decision};
\node (aloss) [as, below=1cm of aact] {$\ell = \rho$};
\node (alossd) [desc, right=0pt of aloss] {Loss};

% info
\node (public) [publicbox, below=1.25cm of sprivd, xshift=-1.5cm] {};
\node (publicd) [desc, right=0pt of public] {Shared};
\node (private) [privatebox, right=0.5cm of publicd] {};
\node (privated) [desc, right=0pt of private] {Private};
\node (ass) [assbox, below=0.25cm of public] {};
\node (assd) [desc, right=0pt of ass] {Assumed};
\node (imp) [impbox, right=0.5cm of assd] {};
\node (impd) [desc, right=0pt of imp] {Implied};


% Statistician
\draw[thick, -] (sdata) -- (mech);
\draw[thick, -Latex] (mech) -- (sstat);


% Adversary
\draw[thick, -Latex] (adata) -- (apost);
\draw[thick, -Latex] (apost) -- (aact);
\draw[thick, -Latex] (aloss) -- (aact);

%Cross
\draw[thick, -Latex] (mech) -- (apost);
\draw[thick, -Latex] (sstat) -- (apost);
\draw[thick, -Latex] (aact) -- (spriv);

% boxes
\node[greybox, fit=(sdata) (sdatad) (sstat) (sstatd) (spriv) (sprivd)] (statbox) {};
\node[greybox, fit=(adata) (adatad) (apost) (apostd) (aact) (aactd) (aloss) (alossd)] (advbox) {};

\node[anchor=south east, xshift=-0.9cm] at (statbox.north) {Sender};
\node[anchor=south east, xshift=2.35cm] at (advbox.north) {Receiver model + assumptions};

\end{tikzpicture}

```


## Measuring privacy

:::{.callout-note}
## Privacy outcome
$$
\inf_{Q \in \mathcal{Q}_x} \rho(d^{Q_T},x)
$$
:::

- Let $S: \mathcal{P} \times \mathsf{X} \rightarrow \mathbb{R}$ such that $S(Q_T,x) = \rho(d^{Q_T},x)$
- $S$ is a (negatively-orientated) proper scoring rule^[@grunwald2004game]
  - A "privacy score"

## Measuring privacy


:::{.callout-note}
## Privacy outcome

$$
\inf_{Q \in \mathcal{Q}_x} S(Q_T,x)
$$
:::

- Use known proper scoring rules to measure privacy outcome
- Create new proper scoring rules from $\rho$
- For example when $d$ is a PDF/PMF:

:::{.fragment}
$\rho(d,x) = -\log d(x)$ then $S(Q_T,x)=-\log q_T(x)$. 
:::

:::{.fragment}
Without loss of generality, we focus on proper scoring rules. 
:::


<!--
## Detour: Proper scoring rules

Proper scoring rules^[@gneiting2007strictly] provide a convenient (and rich) class of functions to assess privacy!

- Consider space of probability distributions $\mathscr{P}$ on $(\mathsf{X},\mathcal{X})$.

- A scoring rule is a function $S:\mathscr{P} \times \mathsf{X} \rightarrow \mathbb{R}$ 

- Expected score $S(P,Q)$
  - If $P,Q \in \mathscr{P}$ define $S(P,Q) = \mathbb{E}_{z\sim Q} S(P,z)$

## Detour: Proper scoring rules

:::{.callout-note}
## Definition 1: Proper scoring rules^[Negatively-orientated]

A scoring rule $S$ is proper with respect to $\mathscr{P}$ if for all $P,Q \in \mathscr{P}$

$$S(Q,Q) \leq S(P,Q).$$

In addition, it is *strictly* proper if $S(P,P) = S(P,Q) \Leftrightarrow P = Q$.

:::

## Scoring rules and privacy

**Forecasting**

- $S(P,x)$ compares a forecast $P$ to a ground truth $x$

- Propriety encourages truthful reporting of the forecast

**Privacy**

- $S(Q,x)$ compares uncertain adversarial knowledge of the data $Q$ to the true data $x$

- Propriety ensures coherent assessment of the privacy


-->

## Measuring privacy

:::{.callout-note}
## Privacy outcome

$$
\inf_{Q \in \mathcal{Q}_x} S(Q_T,x)
$$
:::

- This is an absolute measure of privacy
- We will look at a relative measure of privacy

## Measuring privacy

:::{.callout-note}
## Relative privacy outcome

$$
\inf_{Q \in \mathcal{Q}_x} \left[ S(Q_T,x) - S(Q,x) \right]
$$
:::

- Considers Receiver's information gain
- Sender's relative privacy change (worst-case)

## Measuring privacy


:::{.callout-note}
## Relative privacy outcome

$$
\inf_{Q \in \mathcal{Q}_x} \left[ S(Q_T,x) - S(Q,x) \right]
$$
:::

But...

- $Q_T$ may be random since $T\sim M(~\cdot\mid x)$

# Persuasive privacy

## A privacy definition

- Let $\mathcal{Q}_x \subset \mathscr{P}(\mathsf{X},\mathcal{X})$

- $M: \mathcal{T}\times \mathsf{X} \rightarrow \mathbb{R}_+$ be a mechanism,

- $S$ be a privacy score, constants $\kappa \geq 0$, $0\leq\delta \ll 1$.

:::{.fragment}
:::{.callout-note}
## Definition: Persuasive Privacy

We say $M$ is $(\mathcal{Q}_x, S, \kappa, \delta)$-PP if

$$\inf_{x\in\mathsf{X}}\inf_{Q\in\mathcal{Q}_x}\mathbb{P}_x\left[S(Q, x) - S(Q_{T}, x) \leq \kappa \right] \geq 1 - \delta,$$

where $\mathbb{P}_x$ is w.r.t. $T \sim M(~\cdot\mid x)$.

:::
:::

## A privacy definition

:::{.callout-note}
## Definition: Persuasive Privacy

We say $M$ is $(\mathcal{Q}_x, S, \kappa, \delta)$-PP if

$$\inf_{x\in\mathsf{X}}\inf_{Q\in\mathcal{Q}_x}\mathbb{P}_x\left[S(Q, x) - S(Q_{T}, x) \leq \kappa \right] \geq 1 - \delta,$$

where $\mathbb{P}_x$ is w.r.t. $T \sim M(~\cdot\mid x)$.

:::

- Change in privacy: $S(Q, x) - S(Q_{T}, x)$
- Maximum allowable privacy change: $\kappa$
- Maximum probability of leakage: $\delta\quad$

## Why "persuasive"?

::: {.fragment}
_Sender chooses a mechanism to persuade Receiver to make decisions having limited impact on privacy._
:::

::: {.fragment}
Similar to Bayesian persuasion [@kamenica2011bayesian]
However,
:::

- asymmetry between Sender and Receiver
- utility functions of Sender and Receiver are related
- Sender assesses decisions (mechanism) robustly: worst-case not expected value

## Properties of Persuasive Privacy

1. **Composition**

2. **Receiver post-processing**

:::{.fragment}
:::{.callout-note}
## Definition: Receiver Post-Processing

A guarantee $\mathrm{D}$ satisfies the receiver post-processing property if $M \in \mathfrak{C}(\mathrm{D})$ implies that $M\otimes K \in \mathfrak{C}(\mathrm{D})$ for all Markov kernels $K$ independent of the data $x$.
:::
:::

3. Do not satisfy (Sender) post-processing

## Returning to DP

Recall the Persuasive Privacy elements $(\mathcal{Q}_x, S, \kappa, \delta)$, and consider

1. Family of alternative hypothesis priors

::: {.fragment}
$$\mathcal{H}_x  = \{ Q \in \mathcal P_2:x^\prime \sim x, Q(\{ x , x^\prime \})= 1 \}$$
:::

2. Log-probability scoring rule 

::: {.fragment}
$$S(P,x) = -\log p(x)$$
:::

## Returning to DP

Alternative hypothesis prior class and log-probability score recover PrDP as instance of Persuasive Privacy

::: {.fragment}
$$\inf_{x\sim x^\prime}\mathbb{P}_{x}\left[ m(x,T) \leq \exp\{\epsilon\} m(x^\prime, T) \right] \geq 1 - \delta$$
::: 

- Pure DP is a special case ($\delta = 0$)
- PrDP implies Approximate DP

## Returning to DP

- Pure, Probabilistic, and approximate DP are notions of privacy relative to change in Receiver's knowledge 
- Can be derived by considering persuasive privacy game
- ADP has (Sender) post-processing property, PrDP **does not** (unless pure DP)
  - Typically a criticism of PrDP
- PrDP does have (Receiver) post-processing property
    - A simple fix: control Receiver's decision by including pre-processed statistic

## Privacy for deterministic mechanisms

- DP can't be defined for deterministic mechanisms
- Persuasive privacy framework flexible enough
- Demonstrate privacy for mean $\bar x$

:::{.fragment}
$$M(~\cdot\mid x) = \delta_{\bar x}(\cdot)$$

:::


## Privacy for deterministic mechanisms

Use class of Gaussian distributions

:::{.fragment}

$$\mathcal{G}_{x}^{r} = \left\{ \mathcal{N}(\mu,\Sigma): \frac{(\bar{x}-\bar{\mu})^2}{\overline{\Sigma}} \leq r_1 , c_\Phi \leq r_2 \left(1- \frac{\sigma_i^2}{\Vert\sigma \Vert_2^2} \right) \right\}$$

for $r_1 > 0$ and $r_2 > 1$. 

:::

- Data posterior after observing $\bar x$ is multivariate normal but degenerate
- Support only on subspace $\{z\in \mathbb{R}^n: \bar z = \bar x\}$

## Privacy for deterministic mechanisms

Use (marginal) Dawid--Sebastiani Score

$$ D_i(Q,x) = \log \sigma^2_i(Q) + \frac{[x_i - \mu_i(Q)]^2}{\sigma^{2}_i(Q)}$$

for each element of data, and consider the worst-case element for privacy.

## Privacy for deterministic mechanisms

:::{.callout-note}
## Proposition 

The average mechanism 
    $M(x,\cdot) = \delta_{\bar{x}}$ satisfies $(\mathcal{I},\mathcal{G}_{x}^{r},r_1+ \log r_2,0)$-PP.
:::

- Possible to have a rigorous guarantee for a deterministic mechanism
- Opens path for privacy guarantees for convergent iterative algorithms
  - e.g. MCMC without privacy perturbations (posterior summary statistics)
  - A "decomposition rule"


## Ongoing work

**Short term**

- Decomposition rule (for deterministic mechanisms)
  - Use noise in MCMC as additional privacy

- Nonparametric prior families 

- Work with Independent Commissioner Against Corruption (SA)

**Medium term**
  
- Privacy-utility trade-off

- Computation (SMC, SBI) versus privacy approximation trade-off



:::{.fragment}
Thank you!
:::

## References

::: {#refs}
:::



