---
title: "Persuasive Privacy"
subtitle: "University of Melbourne, 23^rd^ March 2026"
author: "Joshua J Bon"
institute: "Adelaide University"
bibliography: refs.bib
format:
  revealjs:
    html-math-method: katex
    theme: default
    df-print: paged
    incremental: true 
    css: style.css
    chalkboard: true
editor: source
---

## My research

```{dot}
graph G {
  layout=neato;
  subgraph main {
    node [fixedsize=true, width=0.7, height=0.4];
    SBI -- SMC;
    SMC -- Privacy;
    Privacy -- SBI;
  }
  
}
```

## My research

:::: {.columns}

::: {.column width="60%"}

::: {.nonincremental}

**Statistics**

- Bayesian modelling[^1] [^2]
- Efficient computation[^3] [^4] [^5]
- Approximate models^3^ ^4^

:::

:::

::: {.column width="40%"}

```{dot}
graph G {
  layout=neato;
  subgraph main {
    node [fixedsize=true, width=0.7, height=0.4];
    SBI -- SMC;
  }
  
}
```

:::

::::

[^1]: @oflaherty2021teen _Demography_
[^2]: @bon2019polling _JRSS A_
[^3]: @bon2021dasmc _Stat & Comp_
[^4]: @bon2025cal _JMLR_
[^5]: @bon2025rbsmc _arXiv_

## My research

:::: {.columns}

::: {.column width="60%"}

::: {.nonincremental}

**Privacy**

- Federated & transfer learning[^6] [^7]
- Private (Bayesian) computation[^8]
- Approximate models[^9]

:::

:::

::: {.column width="40%"}

```{dot}
graph G {
  layout=neato;
  subgraph main {
    node [fixedsize=true, width=0.7, height=0.4];
    Privacy -- SMC;
    Privacy -- SBI;
  }
  
}
```

:::

::::

[^6]: @hassan2024federated _arXiv_
[^7]: @bretherton2026principled _Bayesian Analysis_
[^8]: @bon2026priv _arXiv_
[^9]: @bon2020acld _J Official Statistics_


## Today's talk

```{dot}
graph G {
  layout=neato;
  subgraph main {
    node [fixedsize=true, width=0.7, height=0.4];
    SBI;
    SMC;
    Privacy [color="red"];
    SBI -- SMC;
    SMC -- Privacy;
    Privacy -- SBI;
  }
  
}
```

## Today's talk

:::{.nonincremental}

1. Differential privacy
2. Privacy from the ground up
3. Persuasive privacy
4. Differential privacy redux
5. Privacy without noise

:::


See: **Bon**, Bailie, Rousseau & Robert (2026) arXiv 2601.22945 

_Under review at ICML 2026_


# Differential privacy

## DP framework

$$\text{Dataset}~x \overset{\text{release}}{\longrightarrow} \text{statistic}~T$$

- $T \sim M(x,\cdot)$, with some mechanism $M$

::: {.fragment}

For example, let $T = \frac{1}{n}\sum_{i=1}^{n} x_i + Z$ with $Z \sim \mathcal{N}(0,\sigma^2)$

:::

::: {.fragment}
$$M(x,\cdot) = \mathcal{N}(\overline{x}_n,\sigma^2)$$
:::

## DP Intuition

Information about data $x$ leaked through statistic $T$ is controlled by how sensitive the mechanism $M(x,\cdot)$ is to $x$.

::: {.fragment}

Measuring sensitivity? I can take the derivative of $M(x,\cdot)$ w.r.t. $x$!

:::

::: {.fragment}

No!

:::

## DP Definition

Borrow from Lipschitz continuity: $M$ is $(\epsilon)$-DP if

$$d_\text{P}[M(x,\cdot),M(x^\prime,\cdot)] \leq \exp(\epsilon) d_\text{X}[x,x^\prime]$$
for all $(x,x^\prime) \in S$ where $S$ is a set of (data) neighbours.

## Laplace mechanism

A fundamental mechanism for DP adds Laplace noise to a deterministic function of the data to ensure privacy.



## DP Definition

Borrow from Lipschitz continuity: $M$ is $(\epsilon,\delta)$-DP if

$$d_\text{P}[M(x,\cdot),M(x^\prime,\cdot)] \leq \exp(\epsilon) d_\text{X}[x,x^\prime] + \delta$$
for all $(x,x^\prime) \in S$ where $S$ is a set of (data) neighbours.

## Some DP Properties

Focusing on $\epsilon$-DP.

1. **Composition**: If $M_1$ and $M_2$ are $\epsilon$-DP then

::: {.fragment}

$$M_1 \otimes M_2 ~\text{is}~ 2\epsilon\text{-DP}$$

:::

2. **Post-processing**: If $K$ is independent of the data then
 
::: {.fragment}

$$M_1 K ~\text{is}~ \epsilon\text{-DP}$$

:::

## Some DP Properties

Focusing on $\epsilon$-DP.

:::{.nonincremental}

1. **Composition**:

If releasing the mean and variance of a dataset is $\epsilon$-DP respectively, then the joint release is $2\epsilon$-DP.

2. **Post-processing**:

If some mechanism $M$ creates a synthetic dataset that is $\epsilon$-DP then any summary statistic from the synthetic dataset will also be $\epsilon$-DP.

:::


## DP Semantics

- Hypothesis testing: Type I & II error controlled by $(\epsilon, \delta)$
  - [@wasserman2010statistical;@hall2013differential]
- Bayesian evidence
  - Prior-to-posterior ratio controlled
    - [@dwork2006calibrating]
  - Posterior-to-posterior ratio controlled
    - [@kasiviswanathan2014semantics]
- Game theory
  - Controls expected utility 
    - [@zhang2021more]

::: {.fragment}
**... post-hoc explanations of DP**
:::



# Privacy from the ground up

## Overview of statistical data privacy

:::{.callout-tip}
## Sender

:::{.fragment}
Government agency release **small area statistics** on a disease
:::

:::

:::{.callout-warning}
## Receiver

:::{.fragment}
Insurance company **raises premiums** using inferred disease prevalence in a small town
:::

:::

:::{.callout-important}
## Effect on data privacy 

:::{.fragment}
Statistic release $\longrightarrow$ adversarial decisions $\longrightarrow$ privacy effect
:::

:::

# Anatomy of statistical data privacy

## The players

### Sender

- Custodian of data
- Will use mechanism to release information about data

### Receiver

- Makes decision based on revealed information
- Decision affects Sender's privacy

## Privacy function

Sender's privacy measured with a **privacy function**

$$\rho:\mathcal{D} \times \mathsf{X} \rightarrow \mathbb{R}$$

- Receiver's decision $\quad d_i \in \mathcal{D}$

- Data value $\quad x \in \mathsf{X}$

:::{.fragment}

$\rho$ orders preferences of decisions for a given dataset
:::


- If $d_1$ is preferred to $d_2$ then $\rho(d_1,x) > \rho(d_2,x)$


## Example 1: Privacy function

For some fixed $\kappa>0$,

$$\rho(d,x) = \begin{cases}
  0, & \text{if } x \in d, \vert d \vert < \kappa\\
  1, & \text{otherwise},
\end{cases}$$


:::{.nonincremental}

- $x \in \mathsf{X} = \mathbb{R}$

- $d \in \mathcal{D} = \{[a,b]:a,b\in \mathbb{R}, a\leq b\}$
:::


## Example 2: Privacy function

<br>

$$\rho(d,x) = -\log d(x)$$

:::{.nonincremental}

- $x \in \mathsf{X} = \mathbb{R}$

- $d \in \mathcal{D} = \{\text{probability density functions on } \mathbb{R}\}$
:::


## The statistic
 
A statistic $T \in \mathsf{T}$ is output from a mechanism $M$ given $x$

$$T \sim M(~\cdot \mid x)$$

- $M:\mathcal{T} \times \mathsf{X} \rightarrow [0,1]$ for $(\mathsf{T},\mathcal{T})$ measurable space

## The statistic

$M$ can be deterministic or randomised

- Summary statistics $\quad\tau:\mathsf{X}\rightarrow\mathsf{T}$
  - $M(~\cdot \mid x) = \delta_{\tau(x)}(\cdot)$ 

- Noisy summary statistics
  - $M(~\cdot \mid x) = \mathcal{N}(~\cdot \mid  \tau(x), \sigma^2)$
  
- Composition of multiple mechanisms

## The statistic

$M$ can be deterministic or randomised

- Result of optimisation
  - MLE: $\quad M(~\cdot \mid x) = \delta_{\tau(x)}(\cdot)$
  - $\tau(x) = \arg\max_{\theta \in \Theta} \log L(x \mid \theta)$
  
- Monte Carlo output
  - MCMC: $\quad M(\mathrm{d}t_{1:N} \mid x) = \prod_{i=1}^N K(\mathrm{d}t_{i} \mid t_{i-1}, x)$

- Noisy optimisation
  - SGD: $\quad M(\mathrm{d}t_N \mid x) = \int_{\mathsf{T}_{-n}}\prod_{i=1}^N K(\mathrm{d}t_{i} \mid t_{i-1}, x)$

## Transperancy

**Privacy class**: A set of mechanisms that satisfy a given privacy definition.

:::{.fragment}

:::{.callout-note}
## Assumption 1: Transperancy

Sender shares the mechanism $M$ and privacy class $\mathfrak{C}$, for which $M \in \mathfrak{C}$, with Receiver. Further, the
definitions of $M$ and $\mathfrak{C}$ do not depend on the data.

:::

:::

## Receiver

:::{.callout-note}
## Assumption 2: Bayesian adversary
Receiver makes Bayesian decisions
:::

1. Holds a prior distribution over the data $Q$
2. Constructs a posterior distribution $Q_T$ from $(M,T)$
3. Make Bayesian decisions
   - loss function $\ell: \mathcal{D} \times \mathsf{X} \rightarrow \mathbb{R}$

## Anatomy of statistical data privacy

```{r dtp, engine = 'tikz',   fig.align = 'center', cache=TRUE}

\begin{tikzpicture}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{fit}
\usetikzlibrary{shapes}

\tikzstyle{st} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{ad} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{pb} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{pbmech} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{bayes} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{desc} = [rectangle, minimum width=2.5cm, rounded corners, minimum height=0.6cm,text centered, fill=gray!10]
\tikzstyle{greybox} = [draw=gray, rounded corners, inner sep=0.5cm, thick]
\tikzstyle{publicbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=teal!30]
\tikzstyle{privatebox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=purple!30]

% Statistician
\node (sdata) [st] {$x$};
\node (sdatad) [desc, left=0pt of sdata] {Data};
\node (mech) [pbmech, below=0.5cm of sdata] {$M$};
\node (mechd) [desc, left=0pt of mech] {Mechanism};
\node (sstat) [pb, below=0.5cm of mech] {$T$};
\node (sstatd) [desc, left=0pt of sstat] {Statistic};
\node (spriv) [st, below=1cm of sstat] {$\rho(d,x)$};
\node (sprivd) [desc, left=0pt of spriv] {Privacy};

% Adversary
\node (adata) [ad, right=3cm of sdata] {$Q$};
\node (adatad) [desc, right=0pt of adata] {Data prior};
\node (apost) [ad, right=3cm of sstat] {$Q_T$};
\node (apostd) [desc, right=0pt of apost] {Data posterior};
\node (aact) [ad, below=1cm of apost] {$d$};
\node (aactd) [desc, right=0pt of aact] {Decision};
\node (aloss) [ad, below=1cm of aact] {$\ell$};
\node (alossd) [desc, right=0pt of aloss] {Loss};

% info
\node (public) [publicbox, below=1.5cm of sprivd, xshift=-1.5cm] {};
\node (publicd) [desc, right=0pt of public] {Shared};
\node (private) [privatebox, right=0.5cm of publicd] {};
\node (privated) [desc, right=0pt of private] {Private};


% Statistician
\draw[thick, -] (sdata) -- (mech);
\draw[thick, -Latex] (mech) -- (sstat);


% Adversary
\draw[thick, -Latex] (adata) -- (apost);
\draw[thick, -Latex] (apost) -- (aact);
\draw[thick, -Latex] (aloss) -- (aact);

%Cross
\draw[thick, -Latex] (mech) -- (apost);
\draw[thick, -Latex] (sstat) -- (apost);
\draw[thick, -Latex] (aact) -- (spriv);

% boxes
\node[greybox, fit=(sdata) (sdatad) (sstat) (sstatd) (spriv) (sprivd)] (statbox) {};
\node[greybox, fit=(adata) (adatad) (apost) (apostd) (aact) (aactd) (aloss) (alossd)] (advbox) {};

\node[anchor=south east, xshift=-0.9cm] at (statbox.north) {Sender};
\node[anchor=south east] at (advbox.north) {Receiver model};

\end{tikzpicture}

```

## Receiver's optimal decision

From Assumption 1:

:::{.callout-note}
## Receiver's optimal decision
$$
d^{P} \in \arg\inf_{d \in \mathcal{D}} \mathbb{E}_{z \sim P}[ \ell(d,z)]
$$
Prior to release $P = Q$, post-release $P = Q_T$.
:::

::: {.fragment}
To measure privacy: **We need assumptions on $Q$ and $\ell$**
:::

## Unknown adversaries

:::{.callout-note}
## Assumption 2: Adversarial loss function
Receiver has loss function $\ell(d,x) = \rho(d,x)$
:::

- Interpretation 1: Receiver targets privacy

:::{.fragment}
If $d^P_{\ell^\prime}$ is Receiver's optimal decision under $(P,\ell^\prime)$ then
:::

::: {.fragment}
$$\mathbb{E}_{x \sim P}[ \rho(d^{P}_\rho,x)] \leq \mathbb{E}_{x \sim P}[ \rho(d^P_{\ell^\prime},x)]$$
:::


- Interpretation 2: data-averaged **worst-case** for Sender

## Unknown adversaries

:::{.callout-note}
## Assumption 3: Adversarial prior class
Let the data-prior $Q \in \mathcal{Q}_x$
:::

:::{.fragment}
Defines class of posteriors and privacy can be measured: 
:::

:::{.fragment}
:::{.callout-note}
## Privacy outcome
$$
d^{Q_T} \in \arg\inf_{d \in \mathcal{D}} \mathbb{E}_{z \sim Q_T}[ \rho(d,z)]
$$

$$
\inf_{Q \in \mathcal{Q}_x} \rho(d^{Q_T},x)
$$
:::
:::

## Anatomy of statistical data privacy

```{r dtp1, engine = 'tikz',   fig.align = 'center', cache=TRUE}

\begin{tikzpicture}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{fit}
\usetikzlibrary{shapes}

\tikzstyle{st} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{ad} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{pb} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{pbmech} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{bayes} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{desc} = [rectangle, minimum width=2.5cm, rounded corners, minimum height=0.6cm,text centered, fill=gray!10]
\tikzstyle{greybox} = [draw=gray, rounded corners, inner sep=0.5cm, thick]
\tikzstyle{publicbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=teal!30]
\tikzstyle{privatebox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=purple!30]

% Statistician
\node (sdata) [st] {$x$};
\node (sdatad) [desc, left=0pt of sdata] {Data};
\node (mech) [pbmech, below=0.5cm of sdata] {$M$};
\node (mechd) [desc, left=0pt of mech] {Mechanism};
\node (sstat) [pb, below=0.5cm of mech] {$T$};
\node (sstatd) [desc, left=0pt of sstat] {Statistic};
\node (spriv) [st, below=1cm of sstat] {$\rho(d,x)$};
\node (sprivd) [desc, left=0pt of spriv] {Privacy};

% Adversary
\node (adata) [ad, right=3cm of sdata] {$Q$};
\node (adatad) [desc, right=0pt of adata] {Data prior};
\node (apost) [ad, right=3cm of sstat] {$Q_T$};
\node (apostd) [desc, right=0pt of apost] {Data posterior};
\node (aact) [ad, below=1cm of apost] {$d$};
\node (aactd) [desc, right=0pt of aact] {Decision};
\node (aloss) [ad, below=1cm of aact] {$\ell$};
\node (alossd) [desc, right=0pt of aloss] {Loss};

% info
\node (public) [publicbox, below=1.25cm of sprivd, xshift=-1.5cm] {};
\node (publicd) [desc, right=0pt of public] {Shared};
\node (private) [privatebox, right=0.5cm of publicd] {};
\node (privated) [desc, right=0pt of private] {Private};


% Statistician
\draw[thick, -] (sdata) -- (mech);
\draw[thick, -Latex] (mech) -- (sstat);


% Adversary
\draw[thick, -Latex] (adata) -- (apost);
\draw[thick, -Latex] (apost) -- (aact);
\draw[thick, -Latex] (aloss) -- (aact);

%Cross
\draw[thick, -Latex] (mech) -- (apost);
\draw[thick, -Latex] (sstat) -- (apost);
\draw[thick, -Latex] (aact) -- (spriv);

% boxes
\node[greybox, fit=(sdata) (sdatad) (sstat) (sstatd) (spriv) (sprivd)] (statbox) {};
\node[greybox, fit=(adata) (adatad) (apost) (apostd) (aact) (aactd) (aloss) (alossd)] (advbox) {};

\node[anchor=south east, xshift=-0.9cm] at (statbox.north) {Sender};
\node[anchor=south east] at (advbox.north) {Receiver model};

\end{tikzpicture}

```

## Anatomy of statistical data privacy

```{r dtp2, engine = 'tikz',   fig.align = 'center', cache=TRUE}

\begin{tikzpicture}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{fit}
\usetikzlibrary{shapes}

\tikzstyle{st} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{ad} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{pb} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{as} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=orange!30]
\tikzstyle{im} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=blue!30]
\tikzstyle{pbmech} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{bayes} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{desc} = [rectangle, minimum width=2.5cm, rounded corners, minimum height=0.6cm,text centered, fill=gray!10]
\tikzstyle{greybox} = [draw=gray, rounded corners, inner sep=0.5cm, thick]
\tikzstyle{publicbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=teal!30]
\tikzstyle{privatebox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=purple!30]
\tikzstyle{assbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=orange!30]
\tikzstyle{impbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=blue!30]

% Statistician
\node (sdata) [st] {$x$};
\node (sdatad) [desc, left=0pt of sdata] {Data};
\node (mech) [pbmech, below=0.5cm of sdata] {$M$};
\node (mechd) [desc, left=0pt of mech] {Mechanism};
\node (sstat) [pb, below=0.5cm of mech] {$T$};
\node (sstatd) [desc, left=0pt of sstat] {Statistic};
\node (spriv) [st, below=1cm of sstat] {$\rho(d,x)$};
\node (sprivd) [desc, left=0pt of spriv] {Privacy};

% Adversary
\node (adata) [as, right=3cm of sdata] {$Q\in\mathcal{Q}_x$};
\node (adatad) [desc, right=0pt of adata] {Data prior};
\node (apost) [im, right=3cm of sstat] {$Q_T$};
\node (apostd) [desc, right=0pt of apost] {Data posterior};
\node (aact) [im, below=1cm of apost] {$d$};
\node (aactd) [desc, right=0pt of aact] {Decision};
\node (aloss) [as, below=1cm of aact] {$\ell = \rho$};
\node (alossd) [desc, right=0pt of aloss] {Loss};

% info
\node (public) [publicbox, below=1.25cm of sprivd, xshift=-1.5cm] {};
\node (publicd) [desc, right=0pt of public] {Shared};
\node (private) [privatebox, right=0.5cm of publicd] {};
\node (privated) [desc, right=0pt of private] {Private};
\node (ass) [assbox, below=0.25cm of public] {};
\node (assd) [desc, right=0pt of ass] {Assumed};
\node (imp) [impbox, right=0.5cm of assd] {};
\node (impd) [desc, right=0pt of imp] {Implied};


% Statistician
\draw[thick, -] (sdata) -- (mech);
\draw[thick, -Latex] (mech) -- (sstat);


% Adversary
\draw[thick, -Latex] (adata) -- (apost);
\draw[thick, -Latex] (apost) -- (aact);
\draw[thick, -Latex] (aloss) -- (aact);

%Cross
\draw[thick, -Latex] (mech) -- (apost);
\draw[thick, -Latex] (sstat) -- (apost);
\draw[thick, -Latex] (aact) -- (spriv);

% boxes
\node[greybox, fit=(sdata) (sdatad) (sstat) (sstatd) (spriv) (sprivd)] (statbox) {};
\node[greybox, fit=(adata) (adatad) (apost) (apostd) (aact) (aactd) (aloss) (alossd)] (advbox) {};

\node[anchor=south east, xshift=-0.9cm] at (statbox.north) {Sender};
\node[anchor=south east, xshift=2.35cm] at (advbox.north) {Receiver model + assumptions};

\end{tikzpicture}

```

## Privacy guarantees and leakage

Consider a set of "private" mechanisms $\mathcal{M}$.

A privacy guarantee is the communication of $(M,\mathcal{M})$
such that $M \in \mathcal{M}$. For example,

$$M \in \mathcal{M}_{x} = \{M \in \mathcal{M}: M(~\cdot\mid x) \text{ is private}\}$$

- A privacy guarantee needs to be transparent to be trusted
- A guarantee only for $x$ reveals information about $x$
    - Extra privacy loss: $x \in \{z\in \mathsf{X}: M \in \mathcal{M}_{z}\}  \subset \mathsf{X}$

## Privacy guarantees and leakage

:::{.callout-note}
## Assumption 4: Protected set
Sender announces a protected set $\mathsf{E} \subset \mathsf{X}$ where the mechanism satisfies

$$M \in \mathcal{M}_{\mathsf{E}} = \{M \in \mathcal{M}: M(~\cdot\mid x) \text{ is private } \forall x \in \mathsf{E}\}$$
:::

- Information about $x$ obscured by protecting all $x \in \mathsf{E}$
- But Receiver learns $x \in \mathsf{E}$

## Anatomy of statistical data privacy

```{r dtp3, engine = 'tikz',   fig.align = 'center', cache=TRUE}

\begin{tikzpicture}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{fit}
\usetikzlibrary{shapes}

\tikzstyle{st} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{ad} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{pb} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{as} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=orange!30]
\tikzstyle{im} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=blue!30]
\tikzstyle{pbmech} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{bayes} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{desc} = [rectangle, minimum width=2.5cm, rounded corners, minimum height=0.6cm,text centered, fill=gray!10]
\tikzstyle{greybox} = [draw=gray, rounded corners, inner sep=0.5cm, thick]
\tikzstyle{publicbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=teal!30]
\tikzstyle{privatebox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=purple!30]
\tikzstyle{assbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=orange!30]
\tikzstyle{impbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=blue!30]

% Statistician
\node (sdata) [st] {$x$};
\node (sdatad) [desc, left=0pt of sdata] {Data};
\node (mech) [pbmech, below=0.5cm of sdata] {$M$};
\node (mechd) [desc, left=0pt of mech] {Mechanism};
\node (sstat) [pb, below=0.5cm of mech] {$T$};
\node (sstatd) [desc, left=0pt of sstat] {Statistic};
\node (spriv) [st, below=1cm of sstat] {$\rho(d,x)$};
\node (sprivd) [desc, left=0pt of spriv] {Privacy};

% Adversary
\node (adata) [as, right=3cm of sdata] {$Q\in\mathcal{Q}_x$};
\node (adatad) [desc, right=0pt of adata] {Data prior};
\node (apost) [im, right=3cm of sstat] {$Q_T$};
\node (apostd) [desc, right=0pt of apost] {Data posterior};
\node (aact) [im, below=1cm of apost] {$d$};
\node (aactd) [desc, right=0pt of aact] {Decision};
\node (aloss) [as, below=1cm of aact] {$\ell = \rho$};
\node (alossd) [desc, right=0pt of aloss] {Loss};

% info
\node (public) [publicbox, below=1.25cm of sprivd, xshift=-1.5cm] {};
\node (publicd) [desc, right=0pt of public] {Shared};
\node (private) [privatebox, right=0.5cm of publicd] {};
\node (privated) [desc, right=0pt of private] {Private};
\node (ass) [assbox, below=0.25cm of public] {};
\node (assd) [desc, right=0pt of ass] {Assumed};
\node (imp) [impbox, right=0.5cm of assd] {};
\node (impd) [desc, right=0pt of imp] {Implied};


% Statistician
\draw[thick, -] (sdata) -- (mech);
\draw[thick, -Latex] (mech) -- (sstat);


% Adversary
\draw[thick, -Latex] (adata) -- (apost);
\draw[thick, -Latex] (apost) -- (aact);
\draw[thick, -Latex] (aloss) -- (aact);

%Cross
\draw[thick, -Latex] (mech) -- (apost);
\draw[thick, -Latex] (sstat) -- (apost);
\draw[thick, -Latex] (aact) -- (spriv);

% boxes
\node[greybox, fit=(sdata) (sdatad) (sstat) (sstatd) (spriv) (sprivd)] (statbox) {};
\node[greybox, fit=(adata) (adatad) (apost) (apostd) (aact) (aactd) (aloss) (alossd)] (advbox) {};

\node[anchor=south east, xshift=-0.9cm] at (statbox.north) {Sender};
\node[anchor=south east, xshift=2.35cm] at (advbox.north) {Receiver model + assumptions};

\end{tikzpicture}

```


## Anatomy of statistical data privacy

```{r dtp5, engine = 'tikz',   fig.align = 'center', cache=TRUE, engine.opts = list(template = "tikz2pdf.tex")}

\begin{tikzpicture}

\usetikzlibrary{positioning}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{fit}
\usetikzlibrary{shapes}

\tikzstyle{st} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{ad} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=purple!30]
\tikzstyle{pb} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{as} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=orange!30]
\tikzstyle{im} = [rectangle, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=blue!30]
\tikzstyle{pbmech} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black, fill=teal!30]
\tikzstyle{bayes} = [diamond, rounded corners, minimum width=2cm, minimum height=1cm,text centered, draw=black]
\tikzstyle{desc} = [rectangle, minimum width=2.5cm, rounded corners, minimum height=0.6cm,text centered, fill=gray!10]
\tikzstyle{greybox} = [draw=gray, rounded corners, inner sep=0.5cm, thick]
\tikzstyle{publicbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=teal!30]
\tikzstyle{privatebox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=purple!30]
\tikzstyle{assbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=orange!30]
\tikzstyle{impbox} = [rectangle, rounded corners, minimum width=0.5cm, minimum height=0.5cm,text centered, draw=black, fill=blue!30]

% Statistician
\node (sdata) [st] {$x$};
\node (sdatad) [desc, left=0pt of sdata] {Data};
\node (mech) [pbmech, below=0.5cm of sdata] {$M$};
\node (mechd) [desc, left=0pt of mech] {Mechanism};
\node (sstat) [pb, below=0.5cm of mech] {$T$};
\node (sstatd) [desc, left=0pt of sstat] {Statistic};

\node (strans) [pb, below=0.5cm of sstat] {$\mathfrak{C}$};
\node (stransd) [desc, left=0pt of strans] {Privacy class};

\node (spriv) [st, below=1.5cm of strans] {$\rho(d,x)$};
\node (sprivd) [desc, left=0pt of spriv] {Privacy};

% Adversary
\node (adata) [as, right=3cm of sdata] {$Q\in\mathcal{Q}_x$};
\node (adatad) [desc, right=0pt of adata] {Data prior};
\node (apost) [im, right=3cm of sstat] {$Q_T$};
\node (apostd) [desc, right=0pt of apost] {Data posterior};
\node (aact) [im, below=1cm of apost] {$d$};
\node (aactd) [desc, right=0pt of aact] {Decision};
\node (aloss) [as, below=1cm of aact] {$\ell = \rho$};
\node (alossd) [desc, right=0pt of aloss] {Loss};


% Statistician
\draw[thick, -] (sdata) -- (mech);
\draw[thick, -Latex] (mech) -- (sstat);


% Adversary
\draw[thick, -Latex] (adata) -- (apost);
\draw[thick, -Latex] (apost) -- (aact);
\draw[thick, -Latex] (aloss) -- (aact);

%Cross
\draw[thick, -Latex] (mech) -- (apost);
\draw[thick, -Latex] (sstat) -- (apost);
\draw[thick, -Latex] (aact) -- (spriv);
\draw[thick, -Latex] (strans) -- (apost);

% boxes
\node[greybox, fit=(sdata) (sdatad) (sstat) (sstatd) (spriv) (sprivd) (strans) (stransd)] (statbox) {};
\node[greybox, fit=(adata) (adatad) (apost) (apostd) (aact) (aactd) (aloss) (alossd)] (advbox) {};

\node[anchor=south east, xshift=-0.9cm] at (statbox.north) {Sender};
\node[anchor=south east, xshift=2.35cm] at (advbox.north) {Receiver model + assumptions};

\end{tikzpicture}

```

# Measuring privacy

## Measuring privacy

:::{.callout-note}
## Privacy outcome
$$
\inf_{Q \in \mathcal{Q}_x} \rho(d^{Q_T},x),\quad d^{Q_T} \in \arg\inf_{d \in \mathcal{D}} \mathbb{E}_{z \sim Q_T}[ \rho(d,z)]
$$
:::

- Let $S: \mathcal{P} \times \mathsf{X} \rightarrow \mathbb{R}$ such that $S(Q_T,x) = \rho(d^{Q_T},x)$
- $S$ is a (negatively-orientated) proper scoring rule^[@grunwald2004game]
  - A "privacy score"

## Measuring privacy


:::{.callout-note}
## Privacy outcome

$$
\inf_{Q \in \mathcal{Q}_x} S(Q_T,x)
$$
:::

- Use known proper scoring rules to measure privacy outcome
- Create new proper scoring rules from $\rho$
- For example when $d$ is a PDF/PMF:

:::{.fragment}
$\rho(d,x) = -\log d(x)$ then $S(Q_T,x)=-\log q_T(x)$. 
:::

:::{.fragment}
Without loss of generality, we focus on proper scoring rules. 
:::

## Detour: Proper scoring rules

Proper scoring rules^[@gneiting2007strictly] provide a convenient (and rich) class of functions to assess privacy!

- Consider space of probability distributions $\mathscr{P}$ on $(\mathsf{X},\mathcal{X})$.

- A scoring rule is a function $S:\mathscr{P} \times \mathsf{X} \rightarrow \mathbb{R}$ 

- Expected score $S(P,Q)$
  - If $P,Q \in \mathscr{P}$ define $S(P,Q) = \mathbb{E}_{z\sim Q} S(P,z)$

## Detour: Proper scoring rules

:::{.callout-note}
## Definition 1: Proper scoring rules^[Negatively-orientated]

A scoring rule $S$ is proper with respect to $\mathscr{P}$ if for all $P,Q \in \mathscr{P}$

$$S(P,P) \leq S(P,Q).$$

In addition, it is *strictly* proper if $S(P,P) = S(P,Q) \Leftrightarrow P = Q$.

:::

## Scoring rules and privacy

**Forecasting**

- $S(P,x)$ compares a forecast $P$ to a ground truth $x$

- Propriety encourages truthful reporting of the forecast

**Privacy**

- $S(Q,x)$ compares uncertain adversarial knowledge of the data $Q$ to the true data $x$

- Propriety ensures coherent assessment of the privacy

## Measuring privacy

:::{.callout-note}
## Privacy outcome

$$
\inf_{Q \in \mathcal{Q}_x} S(Q_T,x)
$$
:::

- This is an absolute measure of privacy
- We will look at a relative measure of privacy

## Measuring privacy

:::{.callout-note}
## Relative privacy outcome

$$
\inf_{Q \in \mathcal{Q}_x} \left[ S(Q_T,x) - S(Q,x) \right]
$$
:::

- Considers Receiver's information gain
- Sender's relative privacy change (worst-case)

## Measuring privacy


:::{.callout-note}
## Relative privacy outcome

$$
\inf_{Q \in \mathcal{Q}_x} \left[ S(Q_T,x) - S(Q,x) \right]
$$
:::

But...

- $Q_T$ may be random since $T\sim M(~\cdot\mid x)$
- We need to protect all $x \in \mathsf{E}$

# Defining privacy

## A privacy definition

- Let $\mathcal{Q}_x \subset \mathscr{P}(\mathsf{X},\mathcal{X})$ and $\mathsf{E} \subset \mathsf{X}$ be the protected set, 

- $M^\mathsf{E}: \mathcal{T}\times \mathsf{E} \rightarrow \mathbb{R}_+$ be a mechanism,

- $S$ be a privacy score, constants $\kappa \geq 0$, $0\leq\delta \ll 1$.

:::{.fragment}
:::{.callout-note}
## Definition 2: Scoring rule privacy I

We say $M^\mathsf{E}$ is $(\mathcal{Q}_x, S, \kappa, \delta)$-SRP if

$$\inf_{x\in\mathsf{E}}\inf_{Q\in\mathcal{Q}_x}\mathbb{P}_x\left[S(Q_{T}, x) \geq S(Q, x) - \kappa \right] \geq 1 - \delta,$$

where the probability  $\mathbb{P}_x$ is w.r.t. $T \sim M^\mathsf{E}(~\cdot\mid x)$.

:::
:::

## A privacy definition

:::{.callout-note}
## Definition 2: Scoring rule privacy I

We say $M^\mathsf{E}$ is $(\mathcal{Q}_x, S, \kappa, \delta)$-SRP if

$$\inf_{x\in\mathsf{E}}\inf_{Q\in\mathcal{Q}_x}\mathbb{P}_x\left[S(Q_{T}, x) \geq S(Q, x) - \kappa \right] \geq 1 - \delta,$$

where the probability  $\mathbb{P}_x$ is w.r.t. $T \sim M^\mathsf{E}(~\cdot\mid x)$.

:::

- $S(Q, x) - \kappa\quad$ is the minimum required privacy score
- $\delta\quad$ is maximum probability of privacy leakage

## Scoring rule privacy: Example 1

Consider priors with probability on true $x$ and alternative $x^\prime$ 

:::{.fragment}
Let $H$ be the Hamming distance. For some fixed $x \in \mathsf{E} = \mathsf{X}$, 

$$\mathcal{Q}_x = \bigcup_{\substack{x^{\prime} \in \mathsf{X}\\ H(x,x^{\prime}) = 1}} \left\{Q \in \mathscr{P}(\mathsf{X}): Q(\{x\}) + Q(\{x^\prime\})=1\right\}$$
:::
:::{.fragment}
Further, for $p \in (0,1)$, consider the restriction

$$\mathcal{Q}_x^p = \{Q\in\mathcal{Q}_x:Q(\{x\}) \geq p\}$$
:::

## Scoring rule privacy: Example 1

**Scoring rule**

Let $p = \frac{\mathrm{d}P}{\mathrm{d}\mu}$ be the Radon–Nikodym derivative of $P$ w.r.t. some base measure $\mu$, take

$$S(P,x) = -\log p(x)$$


:::{.fragment}
**Mechanism**

Consider random algorithm $M$ with $m = \frac{\mathrm{d}M}{\mathrm{d}\mu^\prime}$

:::

## Scoring rule privacy: Example 1

It is straightforward to show

$$\inf_{x\in\mathsf{X}}\inf_{Q\in\mathcal{Q}_x^p}\mathbb{P}_x\left[S(Q_{T}, x) \geq S(Q, x) - \kappa \right] \geq 1 - \delta,$$

is equivalent to

:::{.fragment}
$$\inf_{\substack{x,x^{\prime}\in\mathsf{X}\\H(x,x^{\prime}) = 1}}\mathbb{P}_x\left[m(T \mid x) \geq e^\epsilon m(T \mid x^{\prime})\right] \leq \delta$$

with $\epsilon = \log (1-p) - \log(e^{-\kappa} - p)$

:::

## Interpretation: Example 1

$$\inf_{\substack{x,x^{\prime}\in\mathsf{X}\\H(x,x^{\prime}) = 1}}\mathbb{P}_x\left[m(T \mid x) \geq e^\epsilon m(T \mid x^{\prime})\right] \leq \delta$$

- This is Probabilistic Differential Privacy (PrDP)^[@guingona2023comparing]
- Scoring rule privacy contains PrDP as a special case
- PrDP $\Rightarrow$ Approximate Differential Privacy (ADP)^[@dwork2006approxdp]

## Interpretation: Example 1

- PrDP and ADP are relative notions of privacy
- Can be derived by considering adversarial decisions
- PrDP and ADP assume a minimum strength of Receiver
  - $\mathcal{Q}_x^p = \{Q\in\mathcal{Q}_x:Q(\{x\}) \geq p\}$
- ADP has post-processing invariance property, PrDP **does not**
- From our theory: post-processing by the 
    - Receiver **will not** change their posterior distribution
    - Sender **will** change the privacy guarantee

## Scoring rule privacy

- Let $S_I = \{S_i\}_{i \in I}$ be a set of privacy scores

:::{.fragment}
:::{.callout-note}
## Definition 3: Scoring rule privacy II

We say $M^\mathsf{E}$ is $(\mathcal{Q}_x, S_I, \kappa, \delta)$-SRP if

$$\inf_{i \in I}\inf_{x\in\mathsf{E}}\inf_{Q\in\mathcal{Q}_x}
\mathbb{P}_x\left[S_i(Q_{T}, x) \geq S_i(Q, x) - \kappa \right] \geq 1 - \delta,$$

where the probability  $\mathbb{P}_x$ is w.r.t. $T \sim M^\mathsf{E}(~\cdot\mid x)$.

:::
:::

- SRP has composition property if the class of priors is closed under $M$-updating (i.e. conjugate).

## Scoring rule privacy: Example 2

- DP can't be defined for deterministic mechanisms
- SRP are flexible enough to handle this
- We will consider the mean of the data $\bar x$

:::{.fragment}
$$M(~\cdot\mid x) = \delta_{\bar x}(\cdot)$$

:::

:::{.fragment}
$$\mathcal{Q}_x = \{\mathcal{N}(\mu, \Sigma): \mu \in \mathbb{R}^n, \Sigma \in \text{PD}(\mathbb{R}^n)\}$$

:::



## Scoring rule privacy: Example 2

- Prior $Q \in \mathcal{Q}_x$ with $\mu,\Sigma$
    - such that $[\Sigma]_{ij} = \sigma_i\sigma_j\rho_{ij}$ 
- Data posterior after observing $\bar x$ is multivariate normal but degenerate
- Support only on subspace $\{z\in \mathbb{R}^n: \bar z = \bar x\}$

## Scoring rule privacy: Example 2

The $i$th marginal data-posterior distribution is 
${Q_{\bar{x},i} =  \mathcal{N}(\tilde{\mu}_i, \tilde{\sigma}_i^2)}$ with

$$\tilde{\mu}_i = \mu_i + \sigma_i\frac{\nu_i}{\nu}(\bar{x} - \bar{\mu} ), \qquad \tilde{\sigma}_i^2 = \sigma_i^2\left[1 - \frac{\nu_i^2}{\nu}\right]$$

- $\nu_i = \frac{1}{n}\sum_{j=1}^n \rho_{ij}\sigma_j$
- $\nu = \frac{1}{n^2}\sum_{j=1}^n\sum_{k=1}^n \rho_{jk}\sigma_j\sigma_k$

## Scoring rule privacy: Example 2

One choice: entropy scoring rule

$$S(P,x) = \log \sigma_P^2 + \frac{(x-\mu_P)^2}{\sigma_P^2}$$

then we can show

:::{.fragment}
$$S(Q_{\bar{x},i},x) - S(Q_i,x) \geq \log\frac{\lambda_n}{\lambda_1}\left(1- \frac{\sigma_i^2}{\Vert\sigma \Vert_2^2} \right)  - \frac{(\bar{x} - \bar{\mu})^2}{\lambda_1 \Vert \sigma \Vert_2^2}$$
:::

## Scoring rule privacy: Example 2

Therefore we can use the privacy loss

:::{.fragment}
$$\kappa_i(\bar{x}, \mu, \Sigma) = \frac{(\bar{x} - \bar{\mu})^2}{\lambda_1 \Vert \sigma \Vert_2^2} - \log\frac{\lambda_n}{\lambda_1}\left(1- \frac{\sigma_i^2}{\Vert\sigma \Vert_2^2} \right)$$

with probability of leakage $\delta = 0$
:::


## Sketch of extenstion to MCMC 

If $T \sim M(\cdot\mid \bar{x})$ then $Q_T = \mathbb{E}(Q_{\bar{z}} \mid T)$ for $\bar{z} \sim \bar{Q}(\cdot\mid T)$ where 

- $T = (\theta_1,\theta_2, \ldots, \theta_N)$ is a chain of MCMC samples
- $\bar{Q}$ the implied data-prior on the average
- $\bar{Q}(\cdot\mid T)$ is the posterior update

:::{.fragment}
Recall that for all $\bar{x} \in \mathbb{R}$, 
$S(Q_{\bar{x},i},x) - S(Q_{i},x) \geq - \kappa$ 
:::

- Use MVT and/or properties of $S(\cdot,x)$ to bound $S(Q_{T,i},x) = S(\mathbb{E}_{T}(Q_{\bar{z},i}),x)$ with above

## Comparison with Differential Privacy

For privacy analysis of iterative algorithms **with DP** it is common place to:

1. Prove DP for one step
2. Use composition rules to scale 1 step to $n$ steps
    - As $n \rightarrow \infty$ no privacy guarantee

:::{.fragment}

In **our approach** to privacy it is possible to:

1. Prove privacy for the limiting case (deterministic!)
2. Decompose limiting case to prove privacy for some finite $n$

:::


## Ongoing work

- More examples
  - Other (vector) summary statistics
  - Nonparametric prior families 

- Formalise iterative schemes $n \rightarrow \infty$
  - MCMC
  - SGD
  
- Privacy-utility trade off

:::{.fragment}
Thank you!
:::

## References

::: {#refs}
:::

# Appendix

# Differential Privacy


## (Approximate) differential privacy

- Let $H(x,x')$ be the Hamming distance between $x$ and $x'$
- Constants $\epsilon \geq 0$ and $\delta \geq 0$

:::{.fragment}
:::{.callout-note}
## Definition A1: Approximate differential privacy

We say $M$ is $(\epsilon,\delta)$-ADP if

$$M(A\mid x) \leq e^\epsilon M(A\mid x') + \delta$$

for all $A \in \mathcal{T}$ and $\forall x,x' \in \mathsf{X}$ such that $H(x,x') \leq 1$.
:::
:::

 - Typically referred to as DP
 
# Composition property

## Composition property of SRP

:::{.callout-note}
## Definition 4: Closed under $M$-updating

Let $T \sim M(~\cdot\mid x)$ for fixed $x \in \mathsf{X}$. 

A family of distributions $\mathcal{Q}_x$ is closed under $M$-updating if for every $Q \in \mathcal{Q}_x$ the posterior distribution conditional on $T$, $Q_T \in \mathcal{Q}_x$ for all $T \in \mathsf{T}$.

:::


## Composition property of SRP

:::{.fragment}
$M_1^\mathsf{E}:\mathcal{T}_1 \times \mathsf{E} \rightarrow [0,1],$ 
:::

- e.g. $M_1^\mathsf{E}(t_1 \mid x) = K(t_1 \mid x)$ where $x \in \mathsf{E}$

:::{.fragment}
$M_2^\mathsf{E}:\mathcal{T}_2 \times (\mathsf{T}_1 \times \mathsf{E}) \rightarrow [0,1]$
:::

- e.g. $M_2^\mathsf{E}(t_2 \mid t_1, x) = K(t_2 \mid t_1, x)$ where $x \in \mathsf{E}$


## Composition property of SRP

:::{.callout-note}
## Theorem 1: Composition of SRP

If $M_1^\mathsf{E}$ and $M_2^\mathsf{E}$ are $(\mathcal{Q}_x, S_I, \kappa_1, \delta_1)$-SRP and $(\mathcal{Q}_x,S_I, \kappa_2, \delta_2)$-SRP for all $t_1 \in \mathsf{T}_1$ respectively and $\mathcal{Q}_x$ is closed under $M_1^\mathsf{E}$-updating for all $x \in \mathsf{E}$ then $M_{1}^\mathsf{E} \otimes M_2^\mathsf{E}$ is $(\mathcal{Q}_x, S_I, \kappa_1 + \kappa_2, \delta_1 + \delta_2)$-SRP.

:::


[Proof of Theorem @thm-comp-proof]

- Useful for iterative algorithms: e.g. MCMC, Federated SGD

## Proof of composition property

:::{#thm-comp-proof}
## Composition property

:::{.fragment}
If $\mathcal{Q}_x$ is closed under $M_1$-updating then $Q_{T_1} \in \mathcal{Q}_x$ for all $T_1 \in \mathsf{T}_1$ and priors $Q \in \mathcal{Q}_x$.
:::

:::{.fragment}
And if $M_2^\mathsf{E}$ is $(\mathcal{Q}_x, S_I, \kappa_2, \delta_2)$-SRP for all $t_1 \in \mathsf{T}_1$ then
:::

:::{.fragment}
$$\mathbb{P}_{x, t_1}\left[S_a(Q_{t_1,T_2}, x) \geq S_a(Q_{t_1}, x) - \kappa_2 \right] \geq 1 - \delta_2.$$
for all $x\in \mathsf{E}, a\in I, Q\in\mathcal{Q}_x$.
:::
:::

## Proof of composition property

Taking the expectation of 

$$\mathbb{P}_{x, t_1}\left[S_a(Q_{t_1,T_2}, x) \geq S_a(Q_{t_1}, x) - \kappa_2 \right]$$ 

with respect to $M_1(~\cdot \mid x)$ yields

:::{.fragment}
$$\mathbb{P}_x\left[S_a(Q_{T_1,T_2}, x) \geq S_a(Q_{T_1}, x) - \kappa_2 \right] \geq 1 - \delta_2.$$
:::


## Proof of composition property

Finally, consider the requisite probability

$v = \mathbb{P}_x \left[S_a(Q_{T_1,T_2}, x) \geq S_a(Q, x) - \kappa_1 - \kappa_2 \right] \geq$

$\mathbb{P}_x\left[S_a(Q_{T_1,T_2}, x) \geq S_a(Q_{T_1}, x) - \kappa_2 \geq S_a(Q, x) - \kappa_1 - \kappa_2 \right]$

<br>

:::{.fragment}

Then by Boole–Fréchet inequalities

$v \geq \mathbb{P}_x\left[S_a(Q_{T_1,T_2}, x) \geq S_a(Q_{T_1}, x) - \kappa_2 \right]$
$\qquad+  \mathbb{P}_x\left[S_a(Q_{T_1}, x) \geq S_a(Q, x) - \kappa_1 \right] - 1$
:::

:::{.fragment}
$v \geq 1 - (\delta_1 + \delta_2).$
:::




